Efficient FPGA Implementation of\\ Digit Parallel Online Arithmetic Operators

### 2014-05-29 FPT2014/bare_conf.tex
Change description of data

\begin{abstract}
%\boldmath
Online arithmetic has been widely studied for ASIC implementation. Online components are designed to perform computations most significant digit first, resulting in the ability to chain arithmetic operators together for low latency. Recent research also shows that digit parallel online operators can fail more gracefully when operating beyond the deterministic clocking region, in comparison to operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators in the past has required a large area overhead for FPGA implementation. In this paper, we propose novel approaches to implement the key primitives of online arithmetic, adders and multipliers, efficiently on modern Xilinx FPGAs with 6-input LUTs and carry resources. We demonstrate experimentally that in comparison to a direct implementation, the proposed architectures achieve over $67\%$ Slice savings and over $1.2\times$ speed-ups for adders, and more than $3\times$ area savings and $1.5\times$ speed-ups for multipliers.

% significant area savings and frequency speed-ups compared to a direction.

%The performance of the proposed architectures are experimentally demonstrated

%, they are targeting on FPGAs with 4-input LUTs, which are now gradually replaced by 6-input LUTs.

@@

For online arithmetic, two major research directions have been investigated. First, although there has been previous research into FPGA implementations of online arithmetic, this work only focused on online adders; a brief review of relevant works in this area is presented in Section~\ref{Sec:OA_FPGA}. Second, there are existing approaches to build general purpose online adders specifically for FPGAs with 4-input LUTs and with 2 LUTs within a Slice, such as the Altera Cyclone III \cite{AlteraCyclone} and Xilinx Spartan~3 series \cite{XilinxSpartan}. However, as will be discussed in Section~\ref{Sec:OA_FPGA}, these approaches cannot be directly applied on many modern FPGA with 6-input LUTs and 4 LUTs in a Slice, such as the Xilinx Virtex series FPGAs and all Xilinx 7 series FPGAs~\cite{Virtex7}. Moreover, the move to the 6-LUT on the basic architecture opens new opportunities for efficient implementation.
%However, the implementations of other key arithmetic primitives, such as online multipliers, lack exploration.

In this paper, both issues will be addressed. In Section~\ref{Sec:OA_FPGA}, we propose a novel approach to map the digit parallel online adders onto FPGAs with 6-LUTs. The mapping is based on the fast carry logic, and the available resources within a slice can be fully utilized. In Section~\ref{Sec:OM_FPGA}, we optimize the online multiplication algorithm to yield an efficient FPGA implementation of a digit parallel online multiplier. The resource usage of both operators can be determined analytically. We demonstrate experimentally on Xilinx Virtex-6 FPGAs that the proposed designs achieve significant area reduction and performance gain over their original implementations. For online adders, our method achieves over $67\%$ Slice savings and $1.2\times$ frequency speed-ups in comparison to a direct implementation. For online multipliers, the LUTs and Slice utilization of the proposed new architecture drops over $2.4\times$ and $3.2\times$, with a frequency speed-up more than $1.5\times$ in comparison to a direct implementation. In summary, the main contributions of this paper are an efficient mapping of online adders and multipliers to modern FPGAs. (to be completed)
% by fully utilizing the fast carry logic for .

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder2.pdf}
    \vspace{-6ex}
	\caption{An $N$-digit binary digit-parallel online adder. Both inputs and outputs are represented using SD representation. ``3:2'' denotes a 3:2 compressor.}
    %\vspace{2ex}
	\label{Fig:Radix2SD_adder}
\end{figure}

@@

\subsection{Binary Online Multiplication}
Multiplication is another key arithmetic operator. Typically, online multiplication is performed in a recursive digit serial manner, as illustrated in Algorithm~\ref{Algorithm:OnlineMult} \cite{Ercegovac_OnlineMult} where both inputs and outputs are $N$-digit numbers as given in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated MSD-first through a selection function $sel()$. For any radix $r$ and chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence \cite{Ercegovac_OnlineMult}. As radix 2 is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case $sel()$ is given by (\ref{Eq:SelFunc_OM}) \cite{Oregon_OnlineNetwork}. Notice that the selection is based on two integer digits and one fractional digit of $W_{[j]}$.

\begin{algorithm}[tbp]
  \caption{Online Multiplication}  \label{Algorithm:OnlineMult}
  \begin{algorithmic}[1]
   \State \textbf{Initialization:}~$X_{[-\delta]}=Y_{[-\delta]}=P_{[-\delta]}=0$\vspace{.5ex}
    \For{$j=-\delta,~-\delta+1,~\cdots,~2N-1$}\vspace{.5ex}
        \State $H_{[j]} \leftarrow r^{-\delta}\left(x_{j+\delta+1}\cdot Y_{[j+1]}+y_{j+\delta+1}\cdot X_{[j]}\right)$\vspace{.5ex}
        \State $W_{[j]} \leftarrow P_{[j]} + H_{[j]}$\vspace{.5ex}
        \State $z_{j}~ \leftarrow sel(W_{[j]})$\vspace{.5ex}
        \State $P_{[j+1]} \leftarrow r\left(W_{[j]}-Z_{[j]}\right)$\vspace{.5ex}
    \EndFor
  \end{algorithmic}
\end{algorithm}

@@

Specifically in type 1, both works took advantage of the built-in carry resources in FPGAs. Conventionally the ASIC implementation of online adders is based on 4:2 compressors, as outlined in Fig.~\ref{Fig:PreviousWork} within the gray background. However, directly applying this approach in the FPGAs could be less efficient. This is because there is no carry propagation between the 2 FAs within a 4:2 compressor, and the net delay between them can be large. Instead, Kamp et al \cite{FPT09RA} and Hormigo et al \cite{ASAP09RA} described very similar mapping techniques for online adders with SD representations and CS representations, respectively. They share the common idea of using a maps the logic block (\texttt{LB}) within the dotted circle in Fig.~\ref{Fig:PreviousWork}. In this case, the fast-carry logic in the FPGA can be employed, and the delay between the two FAs can be greatly reduced.

@@

However, we notice the major limitations of both approaches that they only target FPGAs with 4-input LUTs (4-LUT) and two LUTs within a logic slice, such as the Xilinx Spartan series and the Altera Cyclone series. This is naturally reasonable because one \texttt{LB} can be mapped to a single Slice. Nevertheless, for FPGAs with 6-input LUTs (6-LUT) and four LUTs in a slice, such as the Xilinx Virtex series and all Xilinx 7 series FPGAs, directly applying these approaches will result in either resource waste or wrong results. For instance, if two \texttt{LB}s are mapped to a slice with four LUTs as seen in Fig.~\ref{Fig:PreWorkVirtex6}, the outputs of \texttt{LB1} should not be used because the its carry input cannot be explicitly initialized. Due to the same reason, the output of \texttt{XORCY2} should not be used.

@@

\subsection{Proposed Mapping Method}
To tackle this problem, we first present an alternative structure of the online adder to enable an efficient FPGA mapping. The structure of a 4-digit online adder is given as an example in Fig.~\ref{Fig:ModifiedOA}. In this equivalent structure, the first FA in each 4:2 compressor is split into 2 parts, which only generate carry and sum respectively. In this case they can be mapped individually on 2 LUTs. The second FA, which generates the outputs, is unchanged and can be implemented using the fast carry logic.

%propose an alternative mapping method that targets on the Slice containing 4 LUTs such as the Xilinx Virtex series and all Xilinx 7 series FPGAs.

\begin{figure*}[tbp]
	\centering
	\includegraphics[width=.85\textwidth]{./Figures/SDadder_FPGA_New_4digit2.pdf}
	\caption{Alternative structure of online adder. Left: an example of 4-digit online adder. The shaded part refers to the 2 logic blocks (LBs) that can be mapped onto 1 Slice. Right: one LB. The dotted boxes outline the logic that can be mapped onto each LUT and the corresponding fast carry logic.}
	\vspace{-1ex}	
	\label{Fig:ModifiedOA}
\end{figure*}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDadder_FastCarry_New2.pdf}
	\caption{Implementation of 2 logic blocks (LBs) in 1 FPGA Slice which contains four 6-LUTs. NC stands for ``Not Care''.}
	%\vspace{-1ex}	
	\label{Fig:SliceNew}
\end{figure}

In this mapping approach, the area figures of the online adder in terms of the number of LUTs and Slices with respect to the operand word-lengths ($N$) are given in (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), respectively.
\begin{eqnarray}\label{Eq:OAareaLUT}
	OA\_LUT = 2N
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:OAareaSlice}
	OA\_Slice = 1+\left\lceil\frac{N-2}{2}\right\rceil
\end{eqnarray}

\subsection{Performance Analysis}\label{subsec:OA_Performance}
In our experiments, we compare the proposed online adder (\texttt{OA\_new}) against the original design which is implemented using functional descriptions in Verilog based on 4:2 compressors (\texttt{OA\_behv}).  In addition, we also include ripple carry adder (\texttt{RCA}) which uses conventional arithmetic, into the comparisons. All designs are targeting on Xilinx Virtex-6 FPGA with speed grade -1. Notice that in order to utilize the carry resources and to avoid logic optimizations by the synthesis tool, the Xilinx primitive component ``\texttt{Carry4}" is used to create \texttt{OA\_new}.

@@

We record the rated frequencies of all designs when varying the operand word-lengths. The results are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing the designs in ISE 14.1. It can be seen that both designs with online arithmetic achieves relatively stable operating frequencies across a variety of operand word-lengths. This is as expected, because the critical path delay in the online adder is theoretically a fixed value independent of the operand precision. In comparison, the rated frequency of the \texttt{RCA} drops dramatically for larger word-lengths. It follows that both online adders outperform \texttt{RCA} for large word-lengths. Furthermore, our approach operates faster than \texttt{OA\_behv}; the speed-ups of \texttt{OA\_new} over \texttt{OA\_behv} are from $1.20\times$ to $1.26\times$.

@@

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_LUTs3.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_Slices3.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different binary adder implementations with respect to a variety of operand word-lengths. (a) Usage of LUTs. (b) Usage of Slices.}
  \label{Fig:AdderArea}
  \vspace{-1ex}
\end{figure}

@@

However, significant area overhead is required in order to implement this type of \texttt{OM}, due to three main reasons. First, in comparison to the conventional multiplier, more iteration stages are needed. For instance in an $N$-digit \texttt{OM}, $(2N+\delta)$ iterations are required to generate $2N$ digits outputs, or $(N+\delta)$ iterations are required to generate the most significant $N$ digits. Second, the logic of each stage is complex as it involves operations such as online addition, digit-vector multiplication (to generate $x_{j+\delta+1}\cdot Y_{[j+1]}$ and $y_{j+\delta+1}\cdot X_{[j]}$ in Algorithm~\ref{Algorithm:OnlineMult}), shifting and other combinational logic blocks such as the selection function. Also, the word-length of signals $H_{[j]}$, $W_{[j]}$ and $P_{[j]}$ is $(N+2+\delta)$, because the selection function needs 1 integer digit and 1 fractional digit to generate $z_j$, as described in Section~\ref{Sec:Background}. Third, this \texttt{OM} architecture is not optimized specifically for FPGA technology.
%In comparison, an $N$-digit array multiplier, there are only $(N-1)$ rows, each of which is basically an $N$-digit ripple-carry adder.

@@

The resource usage of the proposed \texttt{OM} can be calculated as follows. In an $N$-digit \texttt{OM}, in total $(N-1)$ online adders are needed. According to Algorithm~\ref{Algorithm:OM_DigitParallel}, the word-length of each online adder is $(N+2)$-digit (two integer digits and $N$ fractional digits). Therefore based on (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), the number of LUTs and the number of slices used by the \texttt{OM} is given in (\ref{Eq:OMareaLUT}) and (\ref{Eq:OMareaSlice}), respectively. Note that $L_{S1}$ and $S_{S1}$ denote the number of LUTs and Slices used by the first stage, which is built without online adders.

\begin{eqnarray}\label{Eq:OMareaLUT}
	OM\_{LUT} = 2(N+2)(N-1) + {L}_{S1}
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:OMareaSlice}
	OM\_{Slice} = \left(1+\left\lceil{\frac{N}{2}}\right\rceil\right)(N-1)+S_{S1}
\end{eqnarray}

@@

In comparison, the online multiplier offers the flexibility to simplify the structure corresponding to the required precision. This is possible because in an \texttt{OM}, the product digits are generated initially from the MSD, and there is no carry propagation from the LSD to the MSD with the employment of the redundant number system. This will potentially lead to a more area efficient design. For example, the modified structure diagram of a 4-digit \texttt{OM} is illustrated in Fig.~\ref{Fig:PMStructure_MSDhalf}(a). The word-length of signals within each stage can be correspondingly reduced, as shown in Fig.~\ref{Fig:PMStructure_MSDhalf}(b). It can be seen that instead of keeping identical word-length throughout the stages (in Fig.~\ref{Fig:PMStructure}), in the modified structure the signal word-lengths depend on the stage number $j$ and fewer bits of signals are needed for LSD stages.

\subsection{Performance Analysis}
Similar to the experiments as described in Section~\ref{subsec:OA_Performance}, we compare the direct RTL online multiplier (\texttt{OM\_behv}) against the proposed two types of multiplier, which are implemented to generate products with full precision (\texttt{OM\_full}) and half precision (\texttt{OM\_half}), respectively. In addition, we also compare the \texttt{OM} with the conventional binary arithmetic multiplier. The conventional multiplier is created using the Xilinx Core Generator with speed optimization~\cite{XilinxMult}, and it is implemented based on LUTs without DSPs.

The comparison of area in terms of the cost of LUTs and the cost of slices with respect to different operand word-lengths are illustrated in Fig.~\ref{Fig:OMArea}. The results are obtained from ISE 14.1 after mapping the design to the Virtex-6 FPGA. In comparison to \texttt{OM\_initial}, significant area savings have been obtained by the optimized design \texttt{OM\_full}. Specifically, the LUTs and slice utilization drops $2.44\times\sim3.34\times$ and $3.19\times\sim5.45\times$, respectively.

In comparison to the \texttt{CoreGen} multiplier, the area overheads of \texttt{OM\_full} are between $48\%$ and $84\%$ for LUTs and up to $83\%$ for Slices. On the other hand, the \texttt{OM\_half} achieves area reductions for all operand word-lengths when compared against the \texttt{CoreGen} multiplier. The area saving varies from $40\%$ to $55\%$ for LUTs, and $46\%$ to $55\%$ for slices.

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[]{
  \begin{minipage}{0.47\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/OM_LUTslog.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[]{
  \begin{minipage}{0.47\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/OM_Slicelog.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different types of binary multipliers. (a) Usage of LUTs. (b) Usage of Slices}
  \label{Fig:OMArea}
  %\vspace{-2ex}
\end{figure}

We also check the rated frequencies of all four types of multipliers for a variety of operand word-lengths. The frequency values are still obtained from the post place and route timing reports. Notice that input stimulus to the circuits are randomly sampled from a uniform distribution of $N$-digit numbers. The results are plotted in Fig.\ref{Fig:OM_Freq}, from which we can see that in comparison to \texttt{OM\_behv}, our proposed architectures achieve frequency speed-ups varying from $1.49\times$ to $1.98\times$.

We also notice that there is an increasing gap between the proposed online multipliers and the \texttt{CoreGen} multiplier, especially for large operand word-lengths. This is because in the online multiplier, although the carry resources in the FPGA are efficiently used for area reduction, each 4-digit carry logic within a slice is divided into two parts, as shown previously in Section~\ref{Sec:OA_FPGA}. Therefore the critical path delay of an \texttt{OM} is determined by both the carry logic delays, and the net delay among different stages, which is hyperbole for large operand word-lengths. This also explains the fact that \texttt{OM\_full} and \texttt{OM\_half} run at similar frequencies, as the critical path of both designs are identical in theory.

>>>>

\begin{abstract}
%\boldmath
Online arithmetic has been widely studied for ASIC implementation. Online components are designed to perform computations most significant digit first, resulting in the ability to chain arithmetic operators together for low latency. Recent research also shows that digit parallel online operators can fail more gracefully when operating beyond the deterministic clocking region, in comparison to operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators in the past has required a large area overhead for FPGA implementation. In this paper, we propose novel approaches to implement the key primitives of online arithmetic, adders and multipliers, efficiently on modern Xilinx FPGAs with 6-input LUTs and carry resources. We demonstrate experimentally that in comparison to a direct implementation, the proposed architectures achieve over $67\%$ slice savings and over $1.2\times$ speed-ups for adders, and more than $3\times$ area savings and $1.5\times$ speed-ups for multipliers. Furthermore in comparison to standard arithmetic primitives, the area overheads of online operators drop from $8.41\times$ to $1.88\times$ for adders, and $8.11\times$ to $0.54\times$ for multipliers using our approach.

@@

For online arithmetic, two major research directions have been investigated. First, although there has been previous research into FPGA implementations of online arithmetic, this work only focused on online adders; a brief review of relevant works in this area is presented in Section~\ref{Sec:OA_FPGA}. Second, there are existing approaches to build general purpose online adders specifically for FPGAs with 4-input LUTs and with two LUTs within a slice or a logic element (LE), such as the Xilinx Spartan~3 series \cite{XilinxSpartan} and the Altera Cyclone III \cite{AlteraCyclone}. However, as will be discussed in Section~\ref{Sec:OA_FPGA}, these approaches cannot be directly applied on many modern FPGA with 6-input LUTs and four LUTs in a slice, such as the Xilinx Virtex series FPGAs and all Xilinx 7 series FPGAs~\cite{Virtex7}. Moreover, the move to the 6-LUT on the basic architecture opens new opportunities for efficient implementation.
%However, the implementations of other key arithmetic primitives, such as online multipliers, lack exploration.

In this paper, both issues will be addressed. In Section~\ref{Sec:OA_FPGA}, we propose a novel approach to map the digit parallel online adders onto FPGAs with 6-LUTs. The mapping is based on the fast carry logic, and the available resources within a slice can be fully utilized. In Section~\ref{Sec:OM_FPGA}, we optimize the online multiplication algorithm to yield an efficient FPGA implementation of a digit parallel online multiplier. The resource usage of both operators can be determined analytically. We demonstrate experimentally on Xilinx Virtex-6 FPGAs that the proposed designs achieve significant area reduction and performance gain over their original implementations. For online adders, our method achieves over $67\%$ slice savings and $1.2\times$ frequency speed-ups in comparison to a direct implementation. The area overhead drops from up to $8.41\times$ to $1.88\times$ using the proposed online adder in comparison to the standard ripple-carry adder. For online multipliers, the LUTs and slice utilization of the proposed new architecture drops over $2.4\times$ and $3.2\times$, with a frequency speed-up more than $1.5\times$ in comparison to a direct implementation. When compared to the multiplier with standard arithmetic, the direct implementation has an area overhead up to $8.11\times$. This value drops to $1.84\times$ when using the proposed online multiplier. Furthermore, using the proposed structure can even achieve area savings of up to $55\%$ for half precision results in comparison to the standard multiplier.

In summary, the main contributions of this paper are efficient methods to map online adders and multipliers to modern FPGAs. 
% by fully utilizing the fast carry logic for .

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder2.pdf}
    \vspace{-6ex}
	\caption{An $N$-digit binary digit-parallel online adder. Both inputs and outputs are represented using SD representation. ``3:2'' denotes a 3:2 compressor.}
    %\vspace{2ex}
	\label{Fig:Radix2SD_adder}
\end{figure}

@@

\subsection{Binary Online Multiplication}
Multiplication is another key arithmetic operator. Typically, online multiplication is performed in a recursive digit serial manner, as illustrated in Algorithm~\ref{Algorithm:OnlineMult} \cite{Ercegovac_OnlineMult} where both inputs and outputs are $N$-digit numbers as given in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated MSD-first through a selection function $sel()$. For any radix $r$ and chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence. As the binary radix is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case $sel()$ is given by (\ref{Eq:SelFunc_OM}), and the selection is based on two integer digits and one fractional digit of $W_{[j]}$ \cite{Ercegovac_OnlineMult}.
%
\begin{algorithm}[tbp]
  \caption{Online Multiplication}  \label{Algorithm:OnlineMult}
  \begin{algorithmic}[1]
   \State \textbf{Initialization:}~$X_{[-\delta]}=Y_{[-\delta]}=P_{[-\delta]}=0$\vspace{.5ex}
    \For{$j=-\delta,~-\delta+1,~\cdots,~2N-1$}\vspace{.5ex}
        \State $H_{[j]} \leftarrow r^{-\delta}\left(x_{j+\delta+1}\cdot Y_{[j+1]}+y_{j+\delta+1}\cdot X_{[j]}\right)$\vspace{.5ex}
        \State $W_{[j]} \leftarrow P_{[j]} + H_{[j]}$\vspace{.5ex}
        \State $z_{j}~ \leftarrow sel(W_{[j]})$\vspace{.5ex}
        \State $P_{[j+1]} \leftarrow r\left(W_{[j]}-Z_{[j]}\right)$\vspace{.5ex}
    \EndFor
  \end{algorithmic}
\end{algorithm}

@@

Specifically in type 1, both works took advantage of the built-in carry resources in FPGAs. Conventionally the ASIC implementation of online adders is based on 4:2 compressors, as outlined in Fig.~\ref{Fig:PreviousWork} within the gray background. However, directly applying this approach in the FPGAs could be less efficient. This is because there is no carry propagation between the two FAs within a 4:2 compressor, and the net delay between them can be large. Instead, Kamp et al \cite{FPT09RA} and Hormigo et al \cite{ASAP09RA} described very similar mapping techniques for online adders with SD representations and CS representations, respectively. They share the common idea of building the online adder based on the logic block (\texttt{LB}) as highlighted within the dotted circle in Fig.~\ref{Fig:PreviousWork}. In this case, the fast-carry logic in the FPGA can be employed, and the delay between the two 3:2 compressors can be greatly reduced.

@@

However, we notice the major limitations of both approaches that they only target FPGAs with 4-input LUTs (4-LUT) and two LUTs within a logic slice, such as the Xilinx Spartan series and the Altera Cyclone series. This is naturally reasonable because one \texttt{LB} can be mapped to a single slice. Nevertheless, for FPGAs with 6-input LUTs (6-LUT) and four LUTs in a slice, such as the Xilinx Virtex series and all Xilinx 7 series FPGAs, directly applying these approaches will result in either resource waste or wrong results. For instance, if two \texttt{LB}s are mapped to a slice with four LUTs as seen in Fig.~\ref{Fig:PreWorkVirtex6}, the outputs of \texttt{LB1} should not be used because the its carry input cannot be explicitly initialized. Due to the same reason, the output of \texttt{XORCY2} should not be used.

@@

\subsection{Proposed Mapping Method}
To tackle this problem, we first present an alternative structure of the online adder to enable an efficient FPGA mapping. The structure of a 4-digit online adder is given as an example in Fig.~\ref{Fig:ModifiedOA}. In this equivalent structure, the first FA in each 4:2 compressor is split into two parts, which only generate carry and sum respectively. In this case they can be mapped individually on two LUTs. The second FA, which generates the outputs, is unchanged and can be implemented using the fast carry logic.

%propose an alternative mapping method that targets on the slice containing 4 LUTs such as the Xilinx Virtex series and all Xilinx 7 series FPGAs.

\begin{figure*}[tbp]
	\centering
	\includegraphics[width=.85\textwidth]{./Figures/SDadder_FPGA_New_4digit2.pdf}
	\caption{Alternative structure of online adder. Left: an example of 4-digit online adder. The shaded part refers to the 2 logic blocks (LBs) that can be mapped onto 1 slice. Right: one LB. The dotted boxes outline the logic that can be mapped onto each LUT and the corresponding fast carry logic.}
	\vspace{-1ex}	
	\label{Fig:ModifiedOA}
\end{figure*}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDadder_FastCarry_New2.pdf}
	\caption{Implementation of 2 logic blocks (LBs) in 1 FPGA slice which contains four 6-LUTs. NC stands for ``Not Care''.}
	%\vspace{-1ex}	
	\label{Fig:SliceNew}
\end{figure}

In this mapping approach, the area figures of the online adder in terms of the number of LUTs and slices with respect to the operand word-lengths ($N$) are given in (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), respectively.
\begin{eqnarray}\label{Eq:OAareaLUT}
	OA\_LUT = 2N
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:OAareaSlice}
	OA\_slice = 1+\left\lceil\frac{N-2}{2}\right\rceil
\end{eqnarray}

\subsection{Performance Analysis}\label{subsec:OA_Performance}
In our experiments, we compare the proposed online adder (\texttt{OA\_new}) against the original design which is implemented using functional descriptions in Verilog based on 4:2 compressors (\texttt{OA\_behv}).  In addition, we also include ripple carry adder (\texttt{RCA}) which uses conventional arithmetic, into the comparisons. While in our experiments all designs are targeting on Xilinx Virtex-6 FPGA with speed grade -1, our methods can also be applied on other FPGAs with same slice architecture such as the Xilinx 7 series FPGAs. Notice that in order to utilize the carry resources and to avoid logic optimizations by the synthesis tool, the Xilinx primitive component ``\texttt{Carry4}" is used to create \texttt{OA\_new}.

@@

We record the rated frequencies of all designs when varying the operand word-lengths. The results are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing the designs in ISE 14.1. It can be seen that both designs with online arithmetic achieves relatively stable operating frequencies across a variety of operand word-lengths. This is as expected, because the critical path delay in the online adder is theoretically a fixed value independent of the operand precision. In comparison, the rated frequency of the \texttt{RCA} drops dramatically for larger word-lengths. It follows that both online adders outperform \texttt{RCA} for large word-lengths. Specifically, the frequency of \texttt{OA\_behv} is faster than \texttt{RCA} when $N\geqslant30$. In comparison, the proposed online adder is more efficient in performance, as it outperforms \texttt{RCA} for $N>16$. Furthermore, the speed-ups of \texttt{OA\_new} over \texttt{OA\_behv} are from $1.20\times$ to $1.26\times$.

@@

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[Usage of LUTs]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_LUTs3.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[Usage of slices]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_Slices3.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different binary adder implementations with respect to a variety of operand word-lengths.}
  \label{Fig:AdderArea}
  \vspace{-2ex}
\end{figure}

@@

However, significant area overhead is required in order to implement this type of \texttt{OM}, due to three main reasons. First, in comparison to the conventional multiplier, more iteration stages are needed. For instance in an $N$-digit \texttt{OM}, $(2N+\delta)$ iterations are required to generate $2N$ digits outputs, or $(N+\delta)$ iterations are required to generate the most significant $N$ digits. Second, the logic of each stage is complex as it involves operations such as online addition, digit-vector multiplication (to generate $x_{j+\delta+1}\cdot Y_{[j+1]}$ and $y_{j+\delta+1}\cdot X_{[j]}$ in Algorithm~\ref{Algorithm:OnlineMult}), shifting and other combinational logic blocks such as the selection function. Also, the word-length of signals $H_{[j]}$, $W_{[j]}$ and $P_{[j]}$ is $(N+2+\delta)$, because the selection function needs two integer digits and one fractional digit to generate $z_j$, as described in Section~\ref{Sec:Background}. Third, this \texttt{OM} architecture is not optimized specifically for FPGA technology.
%In comparison, an $N$-digit array multiplier, there are only $(N-1)$ rows, each of which is basically an $N$-digit ripple-carry adder.

@@

The resource usage of the proposed \texttt{OM} can be calculated as follows. In an $N$-digit \texttt{OM}, in total $(N-1)$ online adders are needed. According to Algorithm~\ref{Algorithm:OM_DigitParallel}, the word-length of each online adder is $(N+2)$-digit (two integer digits and $N$ fractional digits). Therefore based on (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), the number of LUTs and the number of slices used by the \texttt{OM} is given in (\ref{Eq:OMareaLUT}) and (\ref{Eq:OMareaSlice}), respectively. Note that $L_{S1}$ and $S_{S1}$ denote the number of LUTs and slices used by the first stage, which is built without online adders.

\begin{eqnarray}\label{Eq:OMareaLUT}
	OM\_{LUT} = 2(N+2)(N-1) + {L}_{S1}
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:OMareaSlice}
	OM\_{slice} = \left(1+\left\lceil{\frac{N}{2}}\right\rceil\right)(N-1)+S_{S1}
\end{eqnarray}

@@

In comparison, the \texttt{OM} offers the flexibility to simplify the structure corresponding to the required precision. This is possible because in an \texttt{OM}, the product digits are generated initially from the MSD, and there is no carry propagation from the LSD to the MSD with the employment of the redundant number system. This will potentially lead to a more area efficient design. For example, the modified structure diagram of a 4-digit \texttt{OM} is illustrated in Fig.~\ref{Fig:PMStructure_MSDhalf}(a). The word-length of signals within each stage can be correspondingly reduced, as shown in Fig.~\ref{Fig:PMStructure_MSDhalf}(b). It can be seen that instead of keeping identical word-length throughout the stages (in Fig.~\ref{Fig:PMStructure}), in the modified structure the signal word-lengths depend on the stage number $j$ and fewer bits of signals are needed for LSD stages.

\subsection{Performance Analysis}
Similar to the experiments as described in Section~\ref{subsec:OA_Performance}, we compare the direct RTL online multiplier (\texttt{OM\_behv}) against the proposed two types of multiplier, which are implemented to generate products with full precision (\texttt{OM\_full}) and half precision (\texttt{OM\_half}), respectively. In addition, we also compare the \texttt{OM} with the conventional binary arithmetic multiplier. The conventional multiplier is created using the Xilinx Core Generator with speed optimization~\cite{XilinxMult}, and it is implemented based on LUTs without DSPs for a fairer comparison.

The comparison of area in terms of the cost of LUTs and the cost of slices with respect to different operand word-lengths are illustrated in Fig.~\ref{Fig:OMArea}. The results are obtained from ISE 14.1 after mapping the design to the Virtex-6 FPGA. In comparison to \texttt{OM\_behv}, significant area savings have been obtained by the optimized design \texttt{OM\_full}. Specifically, the LUTs and slice utilization drops $2.44\times\sim3.34\times$ and $3.19\times\sim5.45\times$, respectively.

In comparison to the \texttt{CoreGen} multiplier, the area overhead of online multipliers drops significantly when transferring to the proposed new structure. In order to generate full precision product, the area overheads of \texttt{OM\_behv} against \texttt{CoreGen} are $3.94\times\sim5.82\times$ for LUTs, and $3.58\times\sim8.11\times$ for slices, respectively. Using the proposed structure \texttt{OM\_full} the area overheads drop to $1.48\times\sim1.83\times$ for LUTs and up to $1.84\times$ for slices. Furthermore, if only half precision result is required, \texttt{OM\_half} is more area efficient for all operand word-lengths when compared against the \texttt{CoreGen} multiplier. The area saving varies from $40\%$ to $55\%$ for LUTs, and $46\%$ to $55\%$ for slices.

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[Usage of LUTs]{
  \begin{minipage}{0.47\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/OM_LUTslog.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[Usage of slices]{
  \begin{minipage}{0.47\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/OM_Slicelog.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different types of binary multipliers.}
  \label{Fig:OMArea}
  %\vspace{-2ex}
\end{figure}

We also check the rated frequencies of all multipliers for a variety of operand word-lengths. The frequency values are plotted in Fig.~\ref{Fig:OM_Freq}, which are still obtained from the post place and route timing reports. In Fig.~\ref{Fig:OM_Freq} we see that in comparison to \texttt{OM\_behv}, our proposed architectures achieve frequency speed-ups varying from $1.49\times$ to $1.98\times$.

Notice that there is an increasing gap between the proposed online multipliers and the \texttt{CoreGen} multiplier, especially for large operand word-lengths. This is because in the online multiplier, although the carry resources in the FPGA are efficiently used for area reduction, each 4-digit carry logic within a slice is divided into two parts, as shown previously in Section~\ref{Sec:OA_FPGA}. Therefore the critical path delay of an \texttt{OM} is determined by both the carry logic delays, and the net delay among different stages, which is hyperbole for large operand word-lengths. This also explains the fact that \texttt{OM\_full} and \texttt{OM\_half} run at similar frequencies, as the critical path of both designs are identical in theory.
